% Synchronized to r29821
\section {HTTPD}

\subsection{Additional Settings}

These variables are not present in the configuration and thus have to be 
added to it when needed.

\begin{description}
\config{HTTPD\_USER}{HTTPD\_USER}{HTTPDUSER}
    {By using this option the web server can be run with the rights of 
    another user than ,,root''. This is useful especially if the webserver 
    should display other pages than the admin interface. 
    \achtung {Scripts that need access to configuration files possibly won't 
    run as expected then. Standard scripts will run with any user provided.}}
\end{description}

\subsection{Remarks}

    If TELMOND is installed the telephone numbers of callers will be displayed 
    on the pages 'status' and 'calls'. Names can be assigned via entries in 
    the file opt/etc/phonebook. This file has the same format as IMONC's 
    telephone number file. Phonebooks may be exchanged between IMONC and the 
    router. The format of each row is ``Phone number=Name[,WAV-file]''  
    WAV files will only be used by IMONC and are ignored by the web server.\\

    The complete web interface has been converted to frameless design using css 
    as of version 2.1.12. Really old browsers may have problems displaying this. 
    The advantage is that the look of the pages can be changed almost completely
    simply by editing the css files (mainly /opt/srv/www/css/main.css).\\

    The webserver package was devloped by Thorsten Pohlmann (\email{pohlmann@tetronik.com})
    and is maintained by Tobias Gruetzmacher (\email{fli4l@portfolio16.de}) at the
    moment. The new design (as of version 2.1.12) was realized by
    Helmut Hummel (\email{hh@fli4l.de}).
