% Synchronized to r53804

\chapter{Documentation for Developers}

\section{Common Rules}

In order to include a new package in the OPT database on the fli4l homepage
some rules must be obeyed. Packages that do not comply with these rules may
be removed from the database without warning.

  \begin{enumerate}
    \item \emph{No} file copy actions by the user! fli4l provides a
     sophisticated system to include the fli4l packages into the installation
     archives. All files that should go to the router are located in \texttt{opt/}.

    \item Pack and compress packets properly: packages must be constructed
    in a way to allow effortless unpacking into the fli4l directory.

    \item Packages must be \emph{completely} configurable via the config file.
    Further editing of configuration files shall not be required by the user.
    Keep difficult decisions away from the users and move them to another part
    at the end of the configuration file. \emph{Note: ONLY MODIFY IF YOU KNOW WHAT
    YOU DO}.

    \item Another hint for config files: by its name it has to be seen clearly
    which OPT it belongs to. For example \var{OPT\_\-HTTPD} contains the variables
    \var{OPT\_\-HTTPD}, \var{HTTPD\_\-USER\_\-N}, a.s.o.

    \item Please built binaries as small as possible! If you compile them yourself
    in FBR please remember to deactive any unneeded feature.

    \item Check for Copyrights! If using template files please keep in mind to
    change the Copyright accordingly. This applies also for config-, check- and
    Opt text files. Replace the Copyright with your own name. A documentation
    that was only copied of course has to keep the Copyright of the orginal author!

    \item Please use only free formats as archive types. These are:
    \begin{itemize}
      \item ZIP  (\texttt{.zip})
      \item GZIP (\texttt{.tgz} or \texttt{.tar.gz})
    \end{itemize}
    Please don't use other formats like RAR, ACE, Blackhole, LHA \ldots
    Windows-Installer files (\texttt{.msi}) or self-extracting archives and
    installers (\texttt{.exe}) should \emph{not} be used.
  \end{enumerate}

\marklabel{sec:libc}{
  \section{Compiling Programs}
}

The environment needed for compiling programs is available in the separate
package ``src"'. There it is also documented how to compile own programs for fli4l.

\input {dev_main_modular}

\input {dev_main_scripting}

\input {dev_main_pf}

\input {dev_main_webgui}

\input {dev_main_boot_dial}

\section{Package ``template"'}

To illustrate some of the things described before the fli4l distribution provides the package
``template"'. This explains by small examples how:

\begin{itemize}
\item a configuration files has to look like (\texttt{config/template.txt})
\item a check files is designed (\texttt{check/template.txt})
\item the extended checking mechanisms are used (\texttt{check/template.ext})
\item configuration variables are stored for later use \\(\texttt{opt/etc/rc.d/rc999.template})
\item stored configuration variables are processed \\(\texttt{opt/usr/bin/template\_show\_config})
\end{itemize}



\section{Structure of the Boot Medium}

As of version 1.5 the program \texttt{syslinux} is used for booting.
Its advantage is that a DOS-compatible file system is available on
the boot medium.

The boot medium contains the following files:

\begin{table}[htbp]
\centering
\begin{tabular}[h!]{lp{10cm}}
\texttt{ldlinux.sys}             & the ``boot loader"' \texttt{syslinux} \\
\texttt{syslinux.cfg}            & config file for \texttt{syslinux} \\
\texttt{kernel}                  & Linux kernel\\
\texttt{rootfs.img}              & RootFS: programs needed for booting\\
\texttt{opt.img}                 & Optional files: drivers and packages\\
\texttt{rc.cfg}                  & config file containing the variables from all
files in fli4l's configuration directory\\
\texttt{boot.msg}                & Text for the \texttt{syslinux} boot menu \\
\texttt{boot\_s.msg}             & Text for the \texttt{syslinux} boot menu \\
\texttt{boot\_z.msg}             & Text for the \texttt{syslinux} boot menu \\
\texttt{hd.cfg}                  & config file to assign partitions \\
\end{tabular}
\end{table}

The script \texttt{mkfli4l.sh} (resp. \texttt{mkfli4l.bat}) at first generates the
files \texttt{opt.img}, \texttt{syslinux.cfg} and \texttt{rc.cfg} as well as \texttt{rootfs.img}.
The files needed are determined by the program \var{mkfli4l} (in the \texttt{unix}- resp.
\texttt{windows}-directory). The kernel and other packages are included in those archives.
The file \texttt{rc.cfg} can be found as well in the Opt-archive as on the
boot medium.\footnote{The one in the Opt-archive is needed during early boot,
because no boot volume is mounted at that time.}

Subsequently, the files \texttt{kernel}, \texttt{rootfs.img}, \texttt{opt.img}
and \texttt{rc.cfg} together with the \texttt{syslinux}-files are copied to the disk.

During boot fli4l uses the script \texttt{/etc/rc} to evaluate the file \texttt{rc.cfg} and
integrate the compressed \texttt{opt.img}-archive into the RootFS-RAM-Disk (depending on
the installation type the files are extracted directly into the rootfs ramdisk or integrated
via symbolic links).
Then the scripts in \texttt{/etc/rc.d/} are run in alphanumeric order and thus the drivers
are loaded and all services get started.

\section{Configuration Files}

Here a short list of the files generated  by fli4l
``on-the-fly'' at boot time.

\begin{enumerate}
\item Provider configuration
  \begin{itemize}
  \item         \texttt{etc/ppp/pap-secrets}

  \item         \texttt{etc/ppp/chap-secrets}

  \end{itemize}
\item DNS configuration
  \begin{itemize}
  \item         \texttt{etc/resolv.conf}

  \item         \texttt{etc/dnsmasq.conf}

  \item         \texttt{etc/dnsmasq\_dhcp.conf}

  \item         \texttt{etc/resolv.dnsmasq}

  \end{itemize}


\item Hosts-File
  \begin{itemize}
  \item         \texttt{etc/hosts}
  \end{itemize}



\item imond-configuration
  \begin{itemize}
  \item \texttt{etc/imond.conf}
  \end{itemize}

\end{enumerate}


\subsection{Provider Configuration}

For the providers chosen User-ID and password are adapted in \texttt{etc/ppp/pap-secrets}.

Example for Provider Planet-Interkom:

\begin{example}
\begin{verbatim}
# Secrets for authentication using PAP
# client        server  secret                  IP addresses
"anonymer"      *       "surfer"                *
\end{verbatim}
\end{example}

In this example ``anonymer"' is the USER-ID. As a remote server in principle anybody is
allowed (hence ``*"'). ``surfer"' is the password for the Provider Planet-Interkom.


\subsection{DNS Configuration}


You can use fli4l as a DNS server. Why this is meaningful (and for Windows PCs
in the LAN even mandatory) is explained in the documentation of the ``base"' package.

The resolver file \texttt{etc/resolv.conf} contains the domain name and the
name server to use. It has the following contents (where ``domain.de"' only
is a placeholder for the value of the configuration variable
\var{DOMAIN\_\-NAME}):

\begin{example}
\begin{verbatim}
        search domain.de
        nameserver 127.0.0.1
\end{verbatim}
\end{example}

The DNS server dnsmasq is configured by the file \texttt{etc/dnsmasq.conf}.
It is automatically generated during boot by processing the scripts \texttt{rc040.dns-local}
and \texttt{rc370.dnsmasq} and might look like this:

\begin{example}
\begin{verbatim}
user=dns
group=dns
resolv-file=/etc/resolv.dnsmasq
no-poll
no-negcache
bogus-priv
log-queries
domain-suffix=lan.fli4l
local=/lan.fli4l/
domain-needed
expand-hosts
filterwin2k
conf-file=/etc/dnsmasq_dhcp.conf
\end{verbatim}
\end{example}





\subsection{Hosts File}

    This file contains a mapping of host names to IP addresses. This
    assignment, however, is used only locally on the flil4 and is not visible
    for other computers in the LAN. This file is actually redundant if
    a local DNS server is started in addition.


\subsection{imond Configuration}

The file \texttt{etc/imond.conf} is constructed amongst others from the configuration variables
\var{CIRC\_x\_NAME}, \var{CIRC\_x\_ROUTE}, \var{CIRC\_x\_CHARGEINT}
and \var{CIRC\_x\_TIMES}. It can consist of up to 32 lines
(except for comment lines). Each line has eight columns:

\begin{enumerate}
\item  Range weekday to weekday
\item  Range hour to hour
\item  Device (\texttt{ippp}\emph{X} or \texttt{isdn}\emph{X})
\item  Circuit with default route: ``yes"'/``no"'
\item  Phone number
\item  Name of the circuits
\item  Phone charges per minute in Euros
\item  Charge interval in seconds
\end{enumerate}

    Here an example:

\begin{example}
\begin{verbatim}
#day  hour  device  defroute  phone        name        charge  ch-int
Mo-Fr 18-09 ippp0   yes       010280192306 Addcom      0.0248  60
Sa-Su 00-24 ippp0   yes       010280192306 Addcom      0.0248  60
Mo-Fr 09-18 ippp1   yes       019160       Compuserve  0.019   180
Mo-Fr 09-18 isdn2   no        0221xxxxxxx  Firma       0.08    90
Mo-Fr 18-09 isdn2   no        0221xxxxxxx  Firma       0.03    90
Sa-Su 00-24 isdn2   no        0221xxxxxxx  Firma       0.03    90
\end{verbatim}
\end{example}

Further explanations for Least-Cost-Routing can be found in the documentation of the package ``base"'.
    
\subsection{The File \texttt{/etc/.profile}}

The file \texttt{/etc/.profile} contains user-defined settings for the
shell. To overwrite the default settings you have to create a file \texttt{etc/.profile}
below the configuration directory. You may enter settings for the command prompt or abbreviations
(so-called ``Aliases"') here.

\wichtig{This file may not contain an \texttt{exit}!}

Examples:

\begin{example}
\begin{verbatim}
alias ll='ls -al'
\end{verbatim}
\end{example}

\subsection{\mtr{Scripts in \texttt{/etc/profile.d/}}}

In the directory \texttt{/etc/profile.d/} one may store scripts that will be executed
when starting a shell and thus may influence the shell's environment. 
Typically OPT packages will place scripts there to define special evironment variables
necessary for the programs they contain.

If both scripts in \texttt{/etc/profile.d/} and the
file \texttt{/etc/.profile} exist, scripts in \texttt{/etc/profile.d/}
will be executed \emph{after} the script \texttt{/etc/.profile}.
